---
title: "Assignment 1 - Machine Learning"
author: "Snehitha Anpur"
date: "2022-09-11"
output: pdf_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The dataset considered here is Bank Churn Data downloaded from the below website:
https://www.kaggle.com/datasets/gauravtopre/bank-customer-churn-dataset

```{r}
library(readr)
BankCustomerChurnPrediction <- read_csv("BankChurnData/BankCustomerChurnPrediction.csv")
View(BankCustomerChurnPrediction)
attach(BankCustomerChurnPrediction)
```

Here are the descriptive statistics and of all the variables in the dataset:

```{r}
summary(BankCustomerChurnPrediction)
```

Here are the frequency and proportion tables of categorical variables:

```{r}
table1 <- table(country,gender)
table1
prop.table(table1)
```

Here are the data transformations:
  - log transformation of a variable "age"
  - square root of the variable "balance"

```{r}
BankCustomerChurnPrediction$log_age <- log10(age)
BankCustomerChurnPrediction$sqrt_bal <- sqrt(balance)
head(BankCustomerChurnPrediction)
```

Below are the plots for quantitative variables:
  # Scatter Plot of Age vs Balance
  # Histogram of Credit_Score
  # Density graph of credit_score
  # Density graph of sqrt_bal
  # Density graph of log_age

```{r}
plot(age,balance,col = 'green') # Scatter Plot of Age vs Balance

hist(credit_score,breaks = 20,col = 'blue') # Histogram of Credit_Score

#install.packages("ggplot2") # Install package ggplot2
library(ggplot2) # Initializing ggplot2 library
#install.packages("dplyr") # Install package dplyr
library(dplyr) # Initializing dplyr library

BankCustomerChurnPrediction %>%
  ggplot( aes(x=credit_score)) +
    geom_density(fill="yellow", color="#e7ecef") # Density graph of credit_score

BankCustomerChurnPrediction %>%
  ggplot( aes(x=sqrt_bal)) +
    geom_density(fill="purple", color="#e7ecef") # Density graph of sqrt_bal

BankCustomerChurnPrediction %>%
  ggplot( aes(x=log_age)) +
    geom_density(fill="#56b3a3", color="#e7ecef") # Density graph of log_age
```

Below are the plots for categorical variables:
  # Bar plot of Country
  # Box plot of credit_score over gender
  
```{r}
counts <- table(country)
barplot(counts, main = "Country Distribution", xlab = "Country", col = "pink") # Barplot of Country

ggplot(BankCustomerChurnPrediction, aes(x=gender, y=credit_score)) +  geom_boxplot(fill='orange') # Box plot of credit_score over gender

```