---
title: "Assignment 3"
author: "Snehitha Anpur"
date: "2022-10-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
#install.packages("e1071") 
library(dplyr)
library(e1071) 
library(readr)
library(caret)
library(ISLR)
UniBank=read.csv("UniversalBank.csv")

set.seed(111)

#Partitioning train and test data

traindata_index = createDataPartition(UniBank$Personal.Loan, p=0.6, list= FALSE)
traindata = UniBank[traindata_index,]
Validationdata = UniBank[-traindata_index,]
```
```{r}
#Creating Pivot Table for CreditCard, Online and Personal Loan

table(CreditCard=traindata$CreditCard,Online=traindata$Online,Loan=traindata$Personal.Loan)

```
Probability of the customer accepting the loan given the condition that credit card =1 and online=1

```{r}
B= 54/(54+497)
B
```
```{r}
#Creating Pivot tables for Loan vs Online and Loan vs Credit Card

LoanvsOnline= table(Loan=traindata$Personal.Loan,Online=traindata$Online)
print(LoanvsOnline)
LoanvsCC=table(Loan=traindata$Personal.Loan,CreditCard=traindata$CreditCard)
print(LoanvsCC)


```
```{r}
#P(CC = 1 | Loan = 1)
 
D1 = 87/291

#P(Online = 1 | Loan = 1)

D2 = 180/(180+112)

#P(Loan = 1)

D3 = 292/(2708+292)

#P(CC = 1 | Loan = 0)

D4 = 812/(1896+812)

#P(Online = 1 | Loan = 0)

D5 = 1596/(1112+1596)

#P(Loan = 0)

D6 = 1-(292/(2708+292))

```
Probability of given (Loan = 1 | CC = 1, Online = 1) using Naive Bayes probability
```{r}

E=(D1*D2*D3)/((D1*D2*D3)+(D4*D5*D6))
E

```
The Value obtained in B is 0.09800363 and E is 0.101083. When comparing we see that the value obtained in finding probability by solving naive Bayes problem is more accurate than that obtained in B.



```{r}
library(dplyr)
#Creating the test data for creditcard=1 and online=1

Test= filter(traindata,CreditCard==1 & Online==1)

#Building Naive Bayes model on the training data

Model=naiveBayes(Personal.Loan~Online+CreditCard,data=traindata)

#Predicting the test data using the Naive Bayes model

Predicted_Test_labels <-predict(Model,Test, type = "raw")

head(Predicted_Test_labels)

```
Value obtained in G is 0.1010447 and E is 0.101083 which are almost same
 
 
 



