---
title: "Machine_Learning_Final_Project"
author: "Snehitha Anpur"
date: "2022-12-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Required Libraries

```{r}


library(dplyr)
library(caret)
library(missForest)
library(corrplot)
library(factoextra)
library(fpc)
library(StatMatch)
library(cluster)
library(ggplot2)
library(cowplot)

```

Loading Source Data

```{r}

Energy_Data=read.csv("D:\\MSBA\\rTutorial\\Rtutorial\\fuel_receipts_costs_eia923.csv")

set.seed(1234)

```

Data Cleaning

```{r}

Energy_Data[Energy_Data==""] = NA

Filtered_Energy_Data  = Energy_Data[,(colMeans(is.na(Energy_Data))*100)<50]

Partitioned_EnergyData_Index = createDataPartition(Filtered_Energy_Data$rowid,p=0.02,list = FALSE)
Partitioned_EnergyData = Filtered_Energy_Data[Partitioned_EnergyData_Index,]

colMeans(is.na(Partitioned_EnergyData))*100

Partitioned_EnergyData$report_date <- as.Date(Partitioned_EnergyData$report_date)

Partitioned_EnergyData$report_date <- as.numeric(format(Partitioned_EnergyData$report_date, "%Y"))

Partitioned_Final_EnergyData=Partitioned_EnergyData[,-c(1,6,8,17)]

```

Data Imputation

```{r}

Partitioned_Final_EnergyData$report_date = as.factor(Partitioned_Final_EnergyData$report_date)

Partitioned_Final_EnergyData$contract_type_code = as.factor(Partitioned_Final_EnergyData$contract_type_code)

Partitioned_Final_EnergyData$energy_source_code = as.factor(Partitioned_Final_EnergyData$energy_source_code)

Partitioned_Final_EnergyData$fuel_group_code = as.factor(Partitioned_Final_EnergyData$fuel_group_code)

Partitioned_Final_EnergyData$primary_transportation_mode_code = as.factor(Partitioned_Final_EnergyData$primary_transportation_mode_code)

Partitioned_Final_EnergyData$natural_gas_transport_code = as.factor(Partitioned_Final_EnergyData$natural_gas_transport_code)

ImputedData = missForest(Partitioned_Final_EnergyData)

Imputed_EnergyData = ImputedData$ximp

Imputed_EnergyData$supplier_name = Partitioned_EnergyData$supplier_name

```

Finding Relations for the Numerical variables

```{r}

corrplot(cor(Imputed_EnergyData[,c(6:11)]))

```


 Data Partition for Train and Test
 
```{r}

Train_label = createDataPartition(Imputed_EnergyData$plant_id_eia,p=0.75,list = FALSE)

Train_EnergyData = Imputed_EnergyData[Train_label,]

Test_EnergyData = Imputed_EnergyData[-Train_label,]

```

Relations for the categorical variables


```{r}

a = ggplot(data = Train_EnergyData, aes(x = report_date,fill = fuel_group_code)) +
    geom_bar(position = "fill") + ylab("proportion") + xlab("Year") +
    stat_count(geom = "text", 
             aes(label = stat(count)),
             position=position_fill(vjust=0.5), colour="white")

b = ggplot(data = Train_EnergyData, aes(x = energy_source_code,fill = fuel_group_code)) +
    geom_bar(position = "fill") + ylab("proportion") + 
    stat_count(geom = "text", 
             aes(label = stat(count)),
             position=position_fill(vjust=0.5), colour="white")

c=ggplot(data = Train_EnergyData, aes(x = energy_source_code,fill = contract_type_code)) +
    geom_bar(position = "fill") + ylab("proportion") +
    stat_count(geom = "text", 
             aes(label = stat(count)),
             position=position_fill(vjust=0.5), colour="white")

d = ggplot(data = Train_EnergyData, aes(x = report_date,fill = primary_transportation_mode_code)) + 
    geom_bar(position = "fill") + ylab("proportion") + xlab("Year") + labs( fill="PTC")
    stat_count(geom = "text", 
             aes(label = stat(count)),
             position=position_fill(vjust=0.5), colour="white")

plot_grid(a,b,c,d)

```

Outliers Removal

```{r}

Fuelunits_quartiles = quantile(Train_EnergyData$fuel_received_units, probs=c(.25, .75), na.rm = FALSE)
Fuelunits_IQR = IQR(Train_EnergyData$fuel_received_units)
 
Fuelunits_Lower = Fuelunits_quartiles[1] - 1.5*Fuelunits_IQR
Fuelunits_Upper = Fuelunits_quartiles[2] + 1.5*Fuelunits_IQR 
 
Filtered_no_outlier = subset(Train_EnergyData, Train_EnergyData$fuel_received_units > Fuelunits_Lower & Train_EnergyData$fuel_received_units < Fuelunits_Upper)

Fuelcost_quartiles = quantile(Filtered_no_outlier$fuel_cost_per_mmbtu, probs=c(.25, .75), na.rm = FALSE)
Fuelcost_IQR <- IQR(Filtered_no_outlier$fuel_cost_per_mmbtu)
 
Fuelcost_Lower = Fuelcost_quartiles[1] - 1.5*Fuelcost_IQR
Fuelcost_Upper = Fuelcost_quartiles[2] + 1.5*Fuelcost_IQR 
 
data_no_outlier = subset(Filtered_no_outlier, Filtered_no_outlier$fuel_cost_per_mmbtu > Fuelcost_Lower & Filtered_no_outlier$fuel_cost_per_mmbtu < Fuelcost_Upper)

```

Choosing and Normalising the  selected attributes

```{r}

Cluster_variables=data_no_outlier[,c(6,7,10,11)]

Norm_EnergyData = scale(Cluster_variables)

```

K-Means Clustering 

```{r}

#fviz_nbclust(Norm_EnergyData, kmeans, method = "wss")

fviz_nbclust(Norm_EnergyData, kmeans, method = "silhouette")

Sil_k4 = kmeans(Norm_EnergyData, centers=4,nstart=50)

fviz_cluster(Sil_k4,data=Norm_EnergyData)

```
DB-Scan Clustering

```{r}

dbscan::kNNdistplot(Norm_EnergyData,k=4)

db=fpc::dbscan(Norm_EnergyData,eps= 0.5,MinPts = 4)

fviz_cluster(db,Norm_EnergyData, stand= FALSE, frame=FALSE,goem= "point")

```
Hierarchical Clustering

```{r}

Get_distance= dist(Norm_EnergyData,method="euclidean")

hclustering=hclust(Get_distance,method = "complete")

plot(hclustering,cex=0.9,hang=-8); rect.hclust(hclustering,k=6,border=1.4);abline(h = 6, col = 'red')

```
Choosing the Clustering Algorithm

```{r}

h_cluster = cutree(hclustering, k=6)

Hierarchial_EnergyData = cbind(data_no_outlier,Cluster=as.factor(h_cluster))

Hierarchial_Analysis = Hierarchial_EnergyData %>% group_by(Cluster,energy_source_code,fuel_group_code) %>% summarise(across(c(fuel_received_units,fuel_mmbtu_per_unit,sulfur_content_pct,ash_content_pct,mercury_content_ppm,fuel_cost_per_mmbtu),mean))

Hierarchial_Mean = Hierarchial_EnergyData %>% group_by(Cluster) %>% summarise(across(c(fuel_received_units,fuel_mmbtu_per_unit,sulfur_content_pct,ash_content_pct,mercury_content_ppm,fuel_cost_per_mmbtu),mean))

Hierarchial_Mean

write.csv(Hierarchial_Mean, "HierarchialMean.csv")

write.csv(Hierarchial_Analysis, "Hierarchial.csv")

write.csv(Hierarchial_EnergyData, "Hierarchial-energydata.csv")

```


```{r}

ggplot(data = Hierarchial_EnergyData, aes(x = report_date,fill = Cluster)) +
    geom_bar(position = "fill") + ylab("proportion") +
    stat_count(geom = "text", 
             aes(label = stat(count)),
             position=position_fill(vjust=0.5), colour="white")

```








